%Version 1.02 @home 9:00 30.12.2014 
%Version notes:
%- The main contents of the acknowledgement written
%- 2 ordering suggestions from friends considered and done
%- Long phrases are to be cut down or use of ; added to improve the phrasing flow of the content
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------
% The Acknowledgements page, for thanking everyone THANK YOU FOR YOUR TIME. HAPPY NEW YEAR 2015. ALL THE BEST 
\acknowledgements{\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics
The road which lead to the accomplishment of my undergraduate project and dissertation is a long one and was not taken alone.

Above all, as I am a faithful believer, and I believe that nothing happens without a reason, I wish to thank God for everything, and making all of this possible whichever way it may have happened.

I also wish to acknowledge the following, first I am very grateful to my supervisor, Dr. G. K. Beeharry, for the opportunity he gave me to work on the topic, he has been an inspirational lecturer who encouraged me to pursue in the field of scientific computing, also most importantly I wish to thank him for his guidance, useful suggestions during the course of this project, and for his reviews, remarks and comments during the writing process of the dissertation. I express my thanks to the lecturers of the department of physics who during those 3 years here at the University of Mauritius contributed to build up the physicist I am today.

My special thanks goes to the MPhil student, Mr. Vinand Prayag, who was the first to trust me to work with him on the specific topic of software correlation, he was always there to talk and help when I needed advices, both on my project and my dissertation.

I am also thankful to the technical staff of the Mauritius Radio Telescope and that of the department of physics for their help and guidance during the practical sessions and the implementation of my project.

Extra credits goes to my classmates with whom we struggled with the BSc (Hons) Physics course, specially those of the computing branch who will recognise themselves, where the last two years together were a lot of fun and made my stay at the university a lot nicer.

I would also like to thank those closest to me for their support and my family, who has always been behind me and helped me greatly to continue in the field of physics at a university level.
}

\clearpage % Start a new page
